### 2025-09-03 cluster-analysis.qmd
20250809 - label functions
20250903 - updates to address updated data files

load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_combine.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_actualStats_career.RData")
```

```{r}
load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_combine.RData")
load(file = "./data/player_stats_weekly.RData")
load(file = "./data/player_stats_seasonal.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_actualStats_career.RData")
```

@@

@@

In our data, all of the following models are equivalent—i.e., they result in the same unequal variance model with a 4-cluster solution—but they arrive there in different ways.

>>>>

load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_combine.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/player_stats_seasonal.RData", fsep = ""))
load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_actualStats_player_career.RData")
```

```{r}
load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_combine.RData")
load(file = "./data/player_stats_weekly.RData")
load(file = "./data/player_stats_seasonal.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_actualStats_player_career.RData")
```

We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.

@@

We can perform a cluster analysis using the `mclust::mclustBIC()`, `mclust::mclustICL()`, and `mclust::mclustBootstrapLRT()` functions of the `mclust` package [@R-mclust; @Scrucca2023_packages].

@@

In our data, all of the following models are equivalent—i.e., they result in the same unequal variance model with a 4-cluster solution—but they arrive there in different ways.
We can fit the cluster model to the optimal number of tiers using the `mclust::Mclust()` function.

### 2025-09-03 download-football-data.qmd
20250902 - ESPN QBR
20250903 - updates to address updated data files

# Downloaded Data - Raw
load(file = "./data/nfl_players_raw.RData")
load(file = "./data/nfl_teams_raw.RData")
load(file = "./data/nfl_rosters_raw.RData")
load(file = "./data/nfl_rosters_weekly_raw.RData")
load(file = "./data/nfl_schedules_raw.RData")
load(file = "./data/nfl_standings_raw.RData")
load(file = "./data/nfl_combine_raw.RData")
load(file = "./data/nfl_draftPicks_raw.RData")
load(file = "./data/nfl_draftValues_raw.RData")
load(file = "./data/nfl_depthCharts_raw.RData")
load(file = "./data/nfl_trades_raw.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp_raw.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown_raw.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation_raw.RData", fsep = ""))
load(file = "./data/nfl_actualStats_position_weekly_raw.RData")
load(file = "./data/nfl_actualStats_weekly_raw.RData")
load(file = "./data/nfl_injuries_raw.RData")
load(file = "./data/nfl_snapCounts_raw.RData")
load(file = "./data/nfl_espnQBR_seasonal_raw.RData")
load(file = "./data/nfl_espnQBR_weekly_raw.RData")
load(file = "./data/nfl_nextGenStats_weekly_raw.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal_raw.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly_raw.RData")
load(file = "./data/nfl_playerContracts_raw.RData")
load(file = "./data/nfl_ftnCharting_raw.RData")
load(file = "./data/nfl_playerIDs_raw.RData")
load(file = "./data/nfl_rankings_draft_raw.RData")
load(file = "./data/nfl_rankings_weekly_raw.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly_raw.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp_raw.RData", fsep = ""))
load(file = "./data/players_projections_seasonal_raw.RData")
load(file = "./data/players_projections_weekly_raw.RData")

# Calculated Data - Raw
load(file = "./data/nfl_actualStats_career_raw.RData")
load(file = "./data/nfl_actualStats_seasonal_raw.RData")
load(file = "./data/nfl_actualFantasyPoints_career_raw.RData")
load(file = "./data/nfl_actualFantasyPoints_seasonal_raw.RData")
load(file = "./data/nfl_actualFantasyPoints_weekly_raw.RData")

# Downloaded Data - Processed
load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_teams.RData")
load(file = "./data/nfl_rosters.RData")
load(file = "./data/nfl_rosters_weekly.RData")
load(file = "./data/nfl_schedules.RData")
load(file = "./data/nfl_standings.RData")
load(file = "./data/nfl_combine.RData")
load(file = "./data/nfl_draftPicks.RData")
load(file = "./data/nfl_draftValues.RData")
load(file = "./data/nfl_depthCharts.RData")
load(file = "./data/nfl_trades.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
load(file = "./data/nfl_actualStats_position_weekly.RData")
load(file = "./data/nfl_actualStats_weekly.RData")
load(file = "./data/nfl_injuries.RData")
load(file = "./data/nfl_snapCounts.RData")
load(file = "./data/nfl_espnQBR_seasonal.RData")
load(file = "./data/nfl_espnQBR_weekly.RData")
load(file = "./data/nfl_nextGenStats_weekly.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/nfl_ftnCharting.RData")
load(file = "./data/nfl_playerIDs.RData")
load(file = "./data/nfl_rankings_draft.RData")
load(file = "./data/nfl_rankings_weekly.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))
load(file = "./data/players_projectedPoints_seasonal.RData")
load(file = "./data/players_projectedPoints_weekly.RData")

# Calculated Data - Processed
load(file = "./data/nfl_actualStats_career.RData")
load(file = "./data/nfl_actualStats_seasonal.RData")
load(file = "./data/nfl_actualFantasyPoints_career.RData")
load(file = "./data/nfl_actualFantasyPoints_seasonal.RData")
load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
load(file = "./data/player_stats_weekly.RData")
load(file = "./data/player_stats_seasonal.RData")
```

@@

```{r}
nfl_players <- nfl_players_raw %>% 
  select(gsis_id, everything()) %>% 
  arrange(display_name)
```

@@

```{r}
nfl_depthCharts <- nfl_depthCharts_raw %>% 
  select(gsis_id, season, week, depth_position, everything()) %>% 
  arrange(full_name, gsis_id, season, week, depth_position)
```

@@

@@

```{r}
nfl_participation_raw <- progressr::with_progress(
  nflreadr::load_participation(
    seasons = 2016:2023, # participation data are no longer available after 2023
    include_pbp = TRUE))
```

@@

We calculate historical actual season-by-season statistics in @sec-calculateActualSeasonalStats.

@@

nfl_actualStats_offense_weekly_raw <- progressr::with_progress(
  nflreadr::load_player_stats(
    seasons = TRUE,
    stat_type = "offense"))

nfl_actualStats_defense_weekly_raw <- progressr::with_progress(
  nflreadr::load_player_stats(
    seasons = TRUE,
    stat_type = "defense"))

nfl_actualStats_kicking_weekly_raw <- progressr::with_progress(
  nflreadr::load_player_stats(
    seasons = TRUE,
    stat_type = "kicking"))
```

save(
  nfl_actualStats_offense_weekly_raw, nfl_actualStats_defense_weekly_raw, nfl_actualStats_kicking_weekly_raw,
  file = "./data/nfl_actualStats_position_weekly_raw.RData"
)
```

load(file = "./data/nfl_actualStats_position_weekly_raw.RData")
```

```{r}
nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly_raw %>% 
  rename(team = recent_team) %>% 
  select(player_id, season, week, everything()) %>% 
  arrange(player_display_name, player_id, season, week)

nfl_actualStats_defense_weekly <- nfl_actualStats_defense_weekly_raw %>% 
  select(player_id, season, week, everything()) %>% 
  arrange(player_display_name, player_id, season, week)

nfl_actualStats_kicking_weekly <- nfl_actualStats_kicking_weekly_raw %>% 
  select(player_id, season, week, everything()) %>% 
  arrange(player_display_name, player_id, season, week)
```

@@

```{r}
nfl_actualStats_offense_weekly %>% 
  group_by(player_id, season, week) %>% 
  filter(n() > 1) %>% 
  head()

nfl_actualStats_defense_weekly %>% 
  group_by(player_id, season, week) %>% 
  filter(n() > 1) %>% 
  head()

nfl_actualStats_kicking_weekly %>% 
  group_by(player_id, season, week) %>% 
  filter(n() > 1) %>% 
  head()
```

```{r}
names(nfl_actualStats_offense_weekly) %>% as.data.frame()

names(nfl_actualStats_defense_weekly) %>% as.data.frame()

names(nfl_actualStats_kicking_weekly) %>% as.data.frame()
```

save(
  nfl_actualStats_offense_weekly, nfl_actualStats_defense_weekly, nfl_actualStats_kicking_weekly,
  file = "./data/nfl_actualStats_position_weekly.RData"
)
```

load(file = "./data/nfl_actualStats_position_weekly.RData")
```

@@

@@

First, we can compute the players' season-by-season statistics using the `nflfastR` [@R-nflfastR] package.

@@

```{r}
nfl_stats_variables
```

@@

save(
  nfl_actualStats_career_player, nfl_actualStats_career_player_inclPost,
  file = "./data/nfl_actualStats_career.RData"
)
```

load(file = "./data/nfl_actualStats_career.RData")
```

@@

We already load players' week-by-week statistics [above](#sec-downloadActualStats) using the `nflreadr::load_player_stats()` function from the `nflreadr` [@R-nflreadr] package.
Nevertheless, we could compute players' weekly statistics from the play-by-play data using the `nflfastR::calculate_stats()` function from the `nflfastR` [@R-nflfastR] package:

@@

# Merge with player info
player_age <- dplyr::left_join(
  player_game_birthdate_gameday,
  nfl_players %>% select(-birth_date, -team_abbr, - team_seq),
  by = c("player_id" = "gsis_id"))

@@

```{r}
# Merge player info with seasonal stats
player_stats_seasonal <- dplyr::full_join(
  nfl_actualFantasyPoints_player_seasonal,
  nfl_players %>% select(-position, -position_group, -team_abbr, - team_seq),
  by = c("player_id" = "gsis_id")
)

>>>>

# Downloaded Data - Raw
load(file = "./data/nfl_players_raw.RData")
load(file = "./data/nfl_teams_raw.RData")
load(file = "./data/nfl_rosters_raw.RData")
load(file = "./data/nfl_rosters_weekly_raw.RData")
load(file = "./data/nfl_schedules_raw.RData")
load(file = "./data/nfl_standings_raw.RData")
load(file = "./data/nfl_combine_raw.RData")
load(file = "./data/nfl_draftPicks_raw.RData")
load(file = "./data/nfl_draftValues_raw.RData")
load(file = "./data/nfl_depthCharts_raw.RData")
load(file = "./data/nfl_trades_raw.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp_raw.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown_raw.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation_raw.RData", fsep = ""))
load(file = "./data/nfl_actualStats_player_seasonal_raw.RData")
load(file = "./data/nfl_actualStats_player_weekly_raw.RData")
load(file = "./data/nfl_actualStats_weekly_raw.RData")
load(file = "./data/nfl_injuries_raw.RData")
load(file = "./data/nfl_snapCounts_raw.RData")
load(file = "./data/nfl_espnQBR_seasonal_raw.RData")
load(file = "./data/nfl_espnQBR_weekly_raw.RData")
load(file = "./data/nfl_nextGenStats_weekly_raw.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal_raw.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly_raw.RData")
load(file = "./data/nfl_playerContracts_raw.RData")
load(file = "./data/nfl_ftnCharting_raw.RData")
load(file = "./data/nfl_playerIDs_raw.RData")
load(file = "./data/nfl_rankings_draft_raw.RData")
load(file = "./data/nfl_rankings_weekly_raw.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly_raw.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp_raw.RData", fsep = ""))
load(file = "./data/players_projections_seasonal_raw.RData")
load(file = "./data/players_projections_weekly_raw.RData")

# Calculated Data - Raw
load(file = "./data/nfl_actualStats_seasonal_raw.RData")
#load(file = "./data/nfl_actualStats_player_career_raw.RData")
load(file = "./data/nfl_actualStats_weekly_raw.RData")
load(file = "./data/nfl_actualFantasyPoints_career_raw.RData")
load(file = "./data/nfl_actualFantasyPoints_seasonal_raw.RData")
load(file = "./data/nfl_actualFantasyPoints_weekly_raw.RData")

# Downloaded Data - Processed
load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_teams.RData")
load(file = "./data/nfl_rosters.RData")
load(file = "./data/nfl_rosters_weekly.RData")
load(file = "./data/nfl_schedules.RData")
load(file = "./data/nfl_standings.RData")
load(file = "./data/nfl_combine.RData")
load(file = "./data/nfl_draftPicks.RData")
load(file = "./data/nfl_draftValues.RData")
load(file = "./data/nfl_depthCharts.RData")
load(file = "./data/nfl_trades.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
load(file = "./data/nfl_actualStats_player_seasonal.RData")
load(file = "./data/nfl_actualStats_player_weekly.RData")
load(file = "./data/nfl_actualStats_weekly.RData")
load(file = "./data/nfl_injuries.RData")
load(file = "./data/nfl_snapCounts.RData")
load(file = "./data/nfl_espnQBR_seasonal.RData")
load(file = "./data/nfl_espnQBR_weekly.RData")
load(file = "./data/nfl_nextGenStats_weekly.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/nfl_ftnCharting.RData")
load(file = "./data/nfl_playerIDs.RData")
load(file = "./data/nfl_rankings_draft.RData")
load(file = "./data/nfl_rankings_weekly.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))
load(file = "./data/players_projectedPoints_seasonal.RData")
load(file = "./data/players_projectedPoints_weekly.RData")

# Calculated Data - Processed
load(file = "./data/nfl_actualStats_seasonal.RData")
load(file = "./data/nfl_actualStats_player_career.RData")
load(file = "./data/nfl_actualStats_weekly.RData")
load(file = "./data/nfl_actualFantasyPoints_career.RData")
load(file = "./data/nfl_actualFantasyPoints_seasonal.RData")
load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
load(file = "./data/player_stats_weekly.RData")
load(file = "./data/player_stats_seasonal.RData")
```

@@

```{r}
nfl_players <- nfl_players_raw %>% 
  select(gsis_id, everything()) %>% # NOTE: the data structure changed in 2025: https://github.com/nflverse/nflverse-players?tab=readme-ov-file#whats-new-in-players-v2
  arrange(display_name)
```

@@

```{r}
nfl_depthCharts <- nfl_depthCharts_raw %>% 
  select(gsis_id, season, week, depth_position, everything()) %>% #NOTE: the data structure will change beginning in the 2025 season: https://nflreadr.nflverse.com/articles/dictionary_depth_charts.html
  arrange(full_name, gsis_id, season, week, depth_position)
```

@@

#nfl_trades_raw <- progressr::with_progress(
#  nflreadr::load_trades(seasons = TRUE))

@@

```{r}
nfl_participation_raw <- progressr::with_progress(
  nflreadr::load_participation(
    seasons = 2016:nflreadr::most_recent_season(), # participation data available after each season has ended per changelog for nflreadr v1.5.0 (https://cran.r-project.org/web/packages/nflreadr/news/news.html)
    include_pbp = TRUE))
```

@@

We can download historical season-by-season actual player statistics using the `nflreadr::load_player_stats()` function from the `nflreadr` package [@R-nflreadr].
Nevertheless, we could also calculate historical actual season-by-season statistics, as in @sec-calculateActualSeasonalStats.
@HoPlayerStats provide a Data Dictionary for statistics for offensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats.html>.
@HoPlayerStatsDefense provide a Data Dictionary for statistics for defensive players at the following link: <https://nflreadr.nflverse.com/articles/dictionary_player_stats_def.html>.
You can also access the data dictionary via `R` using the following syntax:

nfl_actualStats_player_seasonal_reg_raw <- progressr::with_progress(
  nflreadr::load_player_stats(
    seasons = TRUE,
    summary_level = "reg"))

nfl_actualStats_player_seasonal_regPost_raw <- progressr::with_progress(
  nflreadr::load_player_stats(
    seasons = TRUE,
    summary_level = "reg+post"))
```

save(
  nfl_actualStats_player_seasonal_reg_raw, nfl_actualStats_player_seasonal_regPost_raw,
  file = "./data/nfl_actualStats_player_seasonal_raw.Rdata"
)
```

load(file = "./data/nfl_actualStats_player_seasonal_raw.Rdata")
```

The `nfl_actualStats_player_seasonal` objects are in `player`-`season` form.
That is, each row should be uniquely identified by the combination of `player_id` and `season`.
Let's rearrange the data accordingly:

```{r}
nfl_actualStats_player_seasonal_reg <- nfl_actualStats_player_seasonal_reg_raw %>% 
  rename(team = recent_team) %>% 
  select(player_id, season, everything()) %>% 
  arrange(player_display_name, player_id, season)

nfl_actualStats_player_seasonal_regPost <- nfl_actualStats_player_seasonal_regPost_raw %>% 
  select(player_id, season, everything()) %>% 
  arrange(player_display_name, player_id, season)
```

```{r}
nfl_actualStats_player_seasonal_reg %>% 
  group_by(player_id, season) %>% 
  filter(n() > 1) %>% 
  head()

nfl_actualStats_player_seasonal_regPost %>% 
  group_by(player_id, season) %>% 
  filter(n() > 1) %>% 
  head()
```

```{r}
names(nfl_actualStats_player_seasonal_reg) %>% as.data.frame()
```

save(
  nfl_actualStats_player_seasonal_reg, nfl_actualStats_player_seasonal_regPost,
  file = "./data/nfl_actualStats_player_seasonal.RData"
)
```

load(file = "./data/nfl_actualStats_player_seasonal.RData")
```

@@

nfl_actualStats_player_weekly_raw <- progressr::with_progress(
  nflreadr::load_player_stats(
    seasons = TRUE,
    summary_level = "week"))
```

save(
  nfl_actualStats_player_weekly_raw,
  file = "./data/nfl_actualStats_player_weekly_raw.RData"
)
```

load(file = "./data/nfl_actualStats_player_weekly_raw.RData")
```

The `nfl_actualStats_weekly` object is in `player`-`season`-`week` form.
That is, each row should be uniquely identified by the combination of `player_id`, `season`, and `week`.
Let's rearrange the data accordingly:

```{r}
nfl_actualStats_player_weekly <- nfl_actualStats_player_weekly_raw %>% 
  select(player_id, season, week, everything()) %>% 
  arrange(player_display_name, player_id, season, week)
```

@@

```{r}
nfl_actualStats_player_weekly %>% 
  group_by(player_id, season, week) %>% 
  filter(n() > 1) %>% 
  head()
```

```{r}
names(nfl_actualStats_player_weekly) %>% as.data.frame()
```

save(
  nfl_actualStats_player_weekly,
  file = "./data/nfl_actualStats_player_weekly.RData"
)
```

load(file = "./data/nfl_actualStats_player_weekly.RData")
```

@@

ESPN provides Quarterbacks Ratings (QBR).
QBR is described by @Katz2016 (archived at <https://perma.cc/4TP6-E57U>).

@@

We already load players' season-by-season statistics [above](#sec-downloadActualSeasonalStats) using the `nflreadr::load_player_stats()` function from the `nflreadr` [@R-nflreadr] package.
Nevertheless, we could compute players' weekly statistics from the play-by-play data using the `nflfastR::calculate_stats()` function from the `nflfastR` [@R-nflfastR] package:

@@

```{r}
nflfastR::nfl_stats_variables
```

@@

save(
  nfl_actualStats_career_player, nfl_actualStats_career_player_inclPost,
  file = "./data/nfl_actualStats_player_career.RData"
)
```

load(file = "./data/nfl_actualStats_player_career.RData")
```

@@

We already load players' week-by-week statistics [above](#sec-downloadActualWeeklyStats) using the `nflreadr::load_player_stats()` function from the `nflreadr` [@R-nflreadr] package.
Nevertheless, we could compute players' weekly statistics from the play-by-play data using the `nflfastR::calculate_stats()` function from the `nflfastR` [@R-nflfastR] package:

@@

# Merge with player info
player_age <- dplyr::left_join(
  player_game_birthdate_gameday,
  nfl_players %>% select(-birth_date, -latest_team),
  by = c("player_id" = "gsis_id"))

@@

```{r}
# Merge player info with seasonal stats
player_stats_seasonal <- dplyr::full_join(
  nfl_actualFantasyPoints_player_seasonal,
  nfl_players %>% select(-position, -position_group, -latest_team),
  by = c("player_id" = "gsis_id")
)

### 2025-09-03 draft.qmd
20250903 - copyediting
20250903 - updates to address updated data files

@@

For [auction drafts](#sec-draftAuction) and daily fantasy sports (DFS) leagues, a key goal is to select the best lineup or team within the given cap.
For instance, in an [auction drafts](#sec-draftAuction), you may have a \$200 cap and your goal is to draft the best team that you can with the \$200.
Using a mixed-integer programming solver, we can identify the lineup with the most projected points given a particular salary cap.
For projected points for a player, we use their average projected points across all available projection sources.
For a player's estimated cost, we use their [Average Auction Value](#sec-fantasyValueADPAAV) (AAV), rounded up to the nearest integer.

>>>>

We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.
The `players_projections_seasonal` object was derived from projected points objects that were created in @sec-fantasyFootballProjections.

@@

For [auction drafts](#sec-draftAuction) and daily fantasy sports (DFS) leagues, a key goal is to select the best lineup or team within the given cap.
For instance, in an [auction draft](#sec-draftAuction), you may have a \$200 cap and your goal is to draft the best team that you can with the \$200.
Using a mixed-integer programming solver, we can identify the lineup with the most projected points given a particular salary cap.
For projected points for a player, we use their average projected points across all available projection sources.
For a player's estimated cost, we use their [Average Auction Value](#sec-fantasyValueADPAAV) (AAV), rounded up to the nearest integer.
